### 2024-05-14 main.tex
vwr

\paragraph{The model.}
The model we consider is the {\em  multispecies stirring process with open boundaries}, 
i.e. the multispecies analogue of the exclusion process as introduced in \cite{zhou2021orthogonal}, 
with additional boundary reservoirs as proposed in \cite{vanicat2017exact}. 
The hydrodynamic limit and fluctuations of the multispecies stirring process on $\mathbb{Z}^d$
has been recently studied in \cite{casini2023density}.
In our set-up, particles can be of $N-1$ types (or species) and they can move
on a generic connected and undirected graph $G=(V,\mathcal{E})$ with vertex set $V$ and
edge set $\mathcal{E}$. The maximal number of particles allowed at each vertex (also called \textit{maximal occupancy}) is denoted by $\nu\in \mathbb{N}$. 
Two particles do interact when they sit on two sites connected by a graph edge and this interaction consists in swapping them. 
Furthermore, reservoirs are attached at each site of the graph, creating and removing particles. For a precise definition of the rates of the process 
we refer to Section~\ref{sec1}. As we will see, the choice of the boundary reservoirs is motivated by the fact that they allow to determine an absorbing dual process, 
regardless of the graph $G$ on which the process is defined and of the maximal occupancy $\nu$ at each site.

A special case is obtained when the graph $G$ is a finite chain, 
with two reservoirs attached to the first and the last sites respectively, 
and the maximal number of particles allowed at each site is fixed to be $\nu=1$.
In this case the multispecies stirring process with open boundaries 
becomes integrable (see \cite{vanicat2017exact}). 
We recall that for the single species case,  the solution of the exclusion process with open boundaries \cite{derrida1993exact,1993JSP....72..277S} has been of crucial importance in our understanding of the structural properties of non-equilibrium state, such as long-range correlations \cite{spohn1983long} and non-local density large deviation functions \cite{derrida2007non,derrida1998exact} and current large deviations \cite{mallick2022exact,bodineau2005current}. One may argue that the results available for the exclusion process are rooted in the algebraic symmetries of the corresponding integrable spin $s=1/2$ Heisenberg chain, see e.g. \cite{schutzManyBody,SSEPReviewRagoucy} for two excellent reviews.
This integrability property can be combined with duality and leads to closed-form expressions for the correlations and the non-equilibrium steady state, as shown in \cite{frassek2020eigenstates,frassek2020duality}.
In this paper we focus on the multispecies stirring process with open boundaries.
We choose to restrict to the symmetric version of the model, as the formulation
of duality in the presence of both asymmetry and boundary driving is substantially
more involved (for recent results on the single species case see
\cite{schuetz2022reverse,barraquand2022markov}).

@@

At each vertex site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $a$-th component $n_{a}^{x}$ {denotes the number of particles of species $a\in \{1,\ldots,N-1\}$, while the component $n_{N}^{x}$ counts the number of holes at site $x$.} On each vertex we {allow a total maximal number of particles $\nu\in \mathbb{N}$}. {In the following we use indices denoted by lowercase letters (for instance $a,b,c,d\in \{1,\ldots,N-1\}$) when only particles are taken into account. Moreover, we introduce indices denoted by uppercase letters (for instance $A,B,C,D\in\{1,\ldots,N\}$)  to also incorporate the holes, that correspond to the index $N$. The link between the two indices is given by 
	a\in \{1,\ldots,N-1\}\qquad \text{and}\qquad A=(a,N)\,.

@@

\begin{equation}
    \begin{split}
        \mathcal{H}_{x,y}^{T}D&=(R_{x}R_{y})\mathcal{H}_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{u(y)}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{u(y)}\right)\mathcal{H}_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=
        D\mathcal{H}_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$.
 Considering {$\mathcal{A},\mathcal{B}\in {gl}(N)$}, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-\mathcal{B})}\mathcal{A}\exp{(\mathcal{B})}=\mathcal{A}-\left[\mathcal{B},\mathcal{A}\right]+\frac{1}{2!}\left[\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]-\frac{1}{3!}\left[\mathcal{B},\left[\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]\right]+\ldots
 \end{equation}
In the following we evaluate this formula for $\mathcal{B}=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and $\mathcal{A}=E_{CD}$  with $C,D\in\{1,\ldots,N\}$. We find  
\begin{align}
	\left[\mathcal{B},\mathcal{A}\right]=\sum_{a=1}^{N-1}\left[E_{aN},E_{CD}\right]=\delta_{CN}\sum_{a=1}^{N-1}E_{aD}-(1-\delta_{DN})E_{CN}
\end{align}
and 
\begin{align}
	\left[\mathcal{B}\left[\mathcal{B},\mathcal{A}\right]\right]=&\left[\sum_{a=1}^{N-1}E_{aN},\delta_{CN}\sum_{b=1}^{N-1}E_{bD}-(1-\delta_{DN})E_{CN}\right]\nonumber\\=&
	\sum_{a=1}^{N-1}\left(\delta_{CN}\sum_{b=1}^{N-1}\left[E_{aN},E_{bD}\right]-(1-\delta_{DN})\left[E_{aN},E_{CN}\right]\right)\nonumber\\=&
	\sum_{a=1}^{N-1}\left(-\delta_{CN}\delta_{aD}\sum_{b=1}^{N-1}E_{bN}-(1-\delta_{DN})\delta_{CN}E_{aN}\right)\nonumber\\=&
	-\left(\delta_{CN}(1-\delta_{DN})+(1-\delta_{DN})\delta_{CN}\right)\sum_{b=1}^{N-1}E_{bN}\nonumber
	\\=&
	-2\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
\end{align}
From the third commutator on we always obtain zero, i.e. $[\mathcal{B},[\mathcal{B},[\mathcal{B},\mathcal{A}]]]=0$. All in all we have that 
 \begin{align}\label{HT-BA}
 	\fra{\exp{\left(%-\sum_{a=1}^{N-1}E_{aN}
 			-E\right)}E_{CD}\exp{\left(
 		E%\sum_{a=1}^{N-1}E_{aN}
 		\right)}}=E_{CD}-\delta_{CN}\sum_{a=1}^{N-1}E_{aD}+(1-\delta_{DN})E_{CN}-\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
 \end{align}

>>>>

\paragraph{The model.}
The model we consider is the {\em  multispecies stirring process with open boundaries}, 
i.e. the multispecies analogue of the exclusion process as introduced in \cite{zhou2021orthogonal}, 
with additional boundary reservoirs as proposed in \cite{vanicat2017exact}. 
The hydrodynamic limit and fluctuations of the multispecies stirring process on $\mathbb{Z}^d$
has been recently studied in \cite{casini2023density}.
In our set-up, particles can be of $N-1$ types (or species) and they can move
on a generic connected and undirected graph $G=(V,\mathcal{E})$ with vertex set $V$ and
edge set $\mathcal{E}$. The maximal number of particles allowed at each vertex (also called \textit{maximal occupancy}) is denoted by $\nu\in \mathbb{N}$. 
Two particles do interact when they sit on two sites connected by a graph edge and this interaction consists in swapping them. 
Furthermore, reservoirs are attached at each site of the graph, creating and removing particles. For a precise definition of the rates of the process 
we refer to Section~\ref{sec1}. As we will see, the choice of the boundary reservoirs is motivated by the fact that they allow to determine an absorbing dual process
\fra{that does not permit  color-changing until absorption at the boundary}, 
regardless of the graph $G$ on which the process is defined and of the maximal occupancy $\nu$ at each site.

A special case is obtained when the graph $G$ is a finite chain, 
with two reservoirs attached to the first and the last sites respectively, 
and the maximal number of particles allowed at each site is fixed to be $\nu=1$.
In this case the multispecies stirring process with open boundaries 
becomes integrable (see \cite{vanicat2017exact}). 
We recall that for the single species case,  the solution of the exclusion process with open boundaries \cite{derrida1993exact,1993JSP....72..277S} has been of crucial importance in our understanding of the structural properties of non-equilibrium state, such as long-range correlations \cite{spohn1983long} and non-local density large deviation functions \cite{derrida2007non,derrida1998exact} and current large deviations \cite{mallick2022exact,bodineau2005current}. One may argue that the results available for the exclusion process are rooted in the algebraic symmetries of the corresponding integrable spin $s=1/2$ Heisenberg chain, see e.g. \cite{schutzManyBody,SSEPReviewRagoucy} for two excellent reviews.
This integrability property can be combined with duality and leads to closed-form expressions for the correlations and the non-equilibrium steady state, as shown in \cite{frassek2020eigenstates,frassek2020duality}.
In this paper we focus on the multispecies stirring process with open boundaries.
We choose to restrict to the symmetric version of the model, as the formulation
of duality in the presence of bvmmetry and boundary driving is substantially
more involved (for recent results on the single species case see
\cite{schuetz2022reverse,barraquand2022markov}).

@@

At each vertex site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $a$-th component $n_{a}^{x}$ {denotes the number of particles of species $a\in \{1,\ldots,N-1\}$, while the component $n_{N}^{x}$ counts the number of holes at site $x$.} On each vertex we {allow a total maximal number of particles $\nu\in \mathbb{N}$}. {In the following we use indices denoted by lowercase letters (for instance $a,b,c,d\in \{1,\ldots,N-1\}$) when only particles are taken into account. Moreover, we introduce indices denoted by uppercase letters (for instance $A,B,C,D\in\{1,\ldots,N\}$)  to also incorporate the holes, that correspond to the index $N$. 
%The link between the two indices is given by 
%\begin{equation}
%	a\in \{1,\ldots,N-1\}\qquad \text{and}\qquad A=(a,N)\,.
%	\end{equation} 

@@

\begin{equation}
    \begin{split}
        \mathcal{H}_{x,y}^{T}D&=(R_{x}R_{y})\mathcal{H}_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{u(y)}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{u(y)}\right)\mathcal{H}_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=
        D\mathcal{H}_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula \footnote{\fra{See \cite{khuan} for an application of the Hadamard formula in the multispecies asymmetric context.}} the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then we  introduce a creation operator acting on an extra-site $u(x)$.
 Considering {$\mathcal{A},\mathcal{B}\in {gl}(N)$}, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-\mathcal{B})}\mathcal{A}\exp{(\mathcal{B})}=\mathcal{A}-\left[\mathcal{B},\mathcal{A}\right]+\frac{1}{2!}\left[\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]-\frac{1}{3!}\left[\mathcal{B},\left[\mathcal{B},\left[\mathcal{B},\mathcal{A}\right]\right]\right]+\ldots
 \end{equation}
In the following we evaluate this formula for $\mathcal{B}=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and $\mathcal{A}=E_{CD}$  with $C,D\in\{1,\ldots,N\}$. We find  
\begin{align}
	\left[\mathcal{B},\mathcal{A}\right]=\sum_{a=1}^{N-1}\left[E_{aN},E_{CD}\right]=\delta_{CN}\sum_{a=1}^{N-1}E_{aD}-(1-\delta_{DN})E_{CN}
\end{align}
and 
\begin{align}
	\left[\mathcal{B}\left[\mathcal{B},\mathcal{A}\right]\right]=&\left[\sum_{a=1}^{N-1}E_{aN},\delta_{CN}\sum_{b=1}^{N-1}E_{bD}-(1-\delta_{DN})E_{CN}\right]\nonumber\\=&
	\sum_{a=1}^{N-1}\left(\delta_{CN}\sum_{b=1}^{N-1}\left[E_{aN},E_{bD}\right]-(1-\delta_{DN})\left[E_{aN},E_{CN}\right]\right)\nonumber\\=&
	\sum_{a=1}^{N-1}\left(-\delta_{CN}\delta_{aD}\sum_{b=1}^{N-1}E_{bN}-(1-\delta_{DN})\delta_{CN}E_{aN}\right)\nonumber\\=&
	-\left(\delta_{CN}(1-\delta_{DN})+(1-\delta_{DN})\delta_{CN}\right)\sum_{b=1}^{N-1}E_{bN}\nonumber
	\\=&
	-2\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
\end{align}
From the third commutator on we always obtain zero, i.e. $[\mathcal{B},[\mathcal{B},[\mathcal{B},\mathcal{A}]]]=0$. All in all we have that 
 \begin{align}\label{HT-BA}
 	\fra{\exp{\left(%-\sum_{a=1}^{N-1}E_{aN}
 			-E\right)}E_{CD}\exp{\left(
 		E%\sum_{a=1}^{N-1}E_{aN}
 		\right)}}=E_{CD}-\delta_{CN}\sum_{a=1}^{N-1}E_{aD}+(1-\delta_{DN})E_{CN}-\delta_{CN}(1-\delta_{DN})\sum_{b=1}^{N-1}E_{bN}\dt
 \end{align}
